\documentclass[12pt]{report}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage[a4paper, width=150mm,top=25mm,bottom=25mm]{geometry}
\pagestyle{fancy}
\setlength{\headheight}{14.49998pt}
\fancyhead[L]{\small\leftmark}
\fancyhead[R]{\small\rightmark}
% \theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{defn}{Definition}
\newtheorem*{rem}{Remark}
\newtheorem{cor}{Corollary}
\newtheorem{prop}{Proposition}


\begin{document}
\section*{Lecture 1}
\textbf{Remarks:} Basic ordering properties are assumed to be true.
\begin{defn}
    Boundedness: The subset $A \in R$ is said to be bounded above if $\exists M$ such that $M > x~~ \forall x \in A$. And it is bounded below if $\exists m$ such that $m < x~~ \forall x \in A$. If $A$ has both then it is called bounded.   
\end{defn}
\begin{defn}
    Least Upper Bound (lub) Axiom: If $A$ is nonempty subset of $R$ and it is bounded above, then $A$ has a least upper bound in $R$.
\end{defn}
\begin{thm}
    If $A$ is nonempty subset in $R$ and it is bounded below, then it has a greatest lower bound in $R$.
\end{thm}
\begin{proof}
    We first create a set $T$ of lower bounds of $A$ $$ T = \{m \mid m < x~~\forall x \in A\}$$ $T$ is non-empty since $A$ is bounded below. Now, we need to prove that there exits a supremum of $T$ which is also a lower bound of $A$.\\
    Since, set $T$ is bounded above by all the elements of set $A$, it should have a least upper bound, say $M$ such that $ M > m~~ \forall m \in T$. Also, every element of $A$ is an upper bound of $T$ hence by definition of supremum, we can say $M \leq x~~ \forall x \in A$ hence $M$ is the lower bound of $A$. This makes it the greatest lower bound.
\end{proof}
\begin{lem}
    Suppose $A \neq \phi$ and s = lub(A) then for any $y$ such that $y < s$, $\exists a \in A$ such that $ y < a \leq s$.
\end{lem}
\begin{proof}
    Suppose for contradiction, $\nexists$ any element $a$ such that $ y < a$. This means that $y \geq a, \forall a \in A$ $\implies$ $y$ is upper bound of set $A$. But $y$ is already less than least upper bound of set $A$. Hence contradiction. \\
    Therefore, $\exists$ $a \in A$ such that $ y < a \leq s$.   
\end{proof} 
\begin{thm}
    Archimedean Property: Given any positive real numbers $x,y~\exists~n \in N$ such that $ nx > y$.
\end{thm}
\begin{proof}
    Let a set $A = \{nx \mid n \in \mathbb{N}\}$. Suppose for contradiction $ nx \leq y$. Then $y$ is the upper bound of the set $A$. \\
    Let a $x > 0$, then $ y - x < y$ hence $y - x$ is not the upper bound of the set $A$. This means that $\exists m \in \mathbb{N}$ such that $y - x < mx$ $\implies$ $y < mx + x$ $\implies$ $y < (m+1)x$ which is impossible since $(m+1)x \in A$ and $y$ is upper bound of the A.\\
    $ nx > y$ is true.
\end{proof}
\begin{thm}
    If $A$ and $B$ are the two non empty bounded subsets of $R$, such that $x \leq y ~~ \forall x \in A$ and $\forall y \in B$ then $sup(A) \leq inf(B)$
\end{thm}
\begin{proof}
    Let $a$ be the supremum of $A$ and $b$ be the infimum of $B$. Therefore, \linebreak $a \geq x~~\forall x \in A$ and $ b \leq y~~ \forall y \in B$. Also, $A$ is bounded above by $B$ and elements of $B$ are the upper bound for $A$. Hence, $a \leq y ~~ \forall y \in B$. This means that $a$ is the lower bound of $B$ and $a$ is $sup(A)$. In other words, $sup(A) \leq inf(B)$. 
\end{proof}
\begin{thm}
    Given any two real number $a,b$ with $a<b$, $\exists$ $\mathbb{Q}$ between $a$ and $b$.
\end{thm}
\begin{proof}
    Since $b-a > 0$. Take two positive number $b-a$ and $1$ $\exists$ n $\in \mathbb{Z}$ such that $n(b-a) > 1$.\\ TBD
\end{proof}
\begin{thm}
    Any monotone increasing sequence of real numbers that is bounded above converges to some real number.
\end{thm}
\begin{proof}
    Let $x_n$ be a monotone increasing sequence in $\mathbb{R}$ that is bounded above hence there exits a $s$ such that $s = lub\{x_n \mid n \in \mathbb{N}\}$ \\
    Suppose $\epsilon > 0$ $\implies s - \epsilon < s$ and $s - \epsilon$ is not the upper bound of the $x_n$.\\ 
    Using lemma 1, we can say that $\exists x_\epsilon \in x_n$ such that $s - \epsilon < x_\epsilon < s$.\\
    Using monotone condition, for some $n_0 \in \mathbb{N}$, we have,
    $$ s - \epsilon < x_n < s < s + \epsilon~~~~\forall n > n_0 \in \mathbb{N}$$ Hence $ |x_n - s| < \epsilon$. $x_n$ converges to $s~~\forall n > n_0.$ 
\end{proof}
\begin{rem}
    Nested Interval theorem $\approx$ lub $\approx$ theorem $4$
\end{rem}
\begin{proof}
    TBD
\end{proof}
\begin{thm}
    Nested Interval theorem: Suppose $\{I_n\}$ is the sequence of closed and bounded non-empty intervals such that $I_1 \supset I_2 \supset I_3$ \dots then:
    \begin{enumerate}
        \item $\bigcap\limits_{n \geq 1} I_n \neq \phi$.
        \item If the sequence of the length of the intervals goes to $0$ then $\bigcap\limits_{n \geq 1} I_n = \{x\}$. 
    \end{enumerate}
\end{thm}

\begin{proof}
    Let $I_n$ be an interval $[a_n, b_n]$ with $a_m < b_n \forall m,n \in \mathbb{N}$ . Then $\forall n \in \mathbb{N}$, $a_n$ is the increasing sequence and $b_n$ is the decreasing sequence. $b_n$ is upper bound of $a_n$ hence, $a_n < inf(b_n)$.\\
    For $b_n$, $a_n$ is the lower bound of $b_n$ i.e. $sup(a_n)< b_n$. If we combine all inequalities, we get 
    $$ a_n \leq sup(a_n) \leq inf(b_n) \leq b_n ~~ \forall n \in \mathbb{N}$$
    Using density theorem, we can say that $\exists$ some $\mathbb{Q}$ between $sup(a_n)$ and $inf(b_n)$.\\
    Hence, $\bigcap\limits_{n \geq 1} I_n \neq \phi$.\\
    Let the length of the interval to be $L = |b_n - a_n|$. Suppose for contradiction, we have two elements in $\bigcap\limits_{n\geq 1}I_n$ instead of one, say $x$ and $y$. \\ The distance between $x$ and $y$ is $|y - x|$. Since, $L \to 0$ hence $\exists n \in \mathbb{N}$ such that for some $n_0 \geq n$, $|L| = |b_{n_0} - a_{n_0}|< \epsilon$ for some $\epsilon > 0$. Since $|L| \to 0$, we can choose $\epsilon$ such that it is smaller than $|y - x|$. Then, if interval contains any one of the point, it can not contain the other.
\end{proof}

\hrule
\section*{Lecture 2}
\begin{defn}
    Decimals representation of Real numbers: Let $z \in \mathbb{R}^+$ be given. Let $n_0$ be the largest integer such that $n_0 \leq z$. Let $n_1$ be the largest integer such that $n_0 + \frac{n_1}{10} \leq z$. As such, say $n_k$ is defined for some $k$. Let $n_{k+1}$ be the largest integer such that $n_0 + \frac{n_1}{10^1} + \frac{n_1}{10^2} + \dots + \frac{n_k}{10^k}+\frac{n_{k+1}}{10^{k+1}} \leq z$. Consider the set of all such finite sums, i.e. the set of all
    $$ z_k = n_0 + \frac{n_1}{10^1} + \frac{n_1}{10^2} + \dots + \frac{n_k}{10^k}+\frac{n_{k+1}}{10^{k+1}} \leq z$$
    This set has a supremum and that is $z$ itself. We symbolically write $z = n_0.n_1n_2 \dots $
\end{defn}
\begin{lem}
    Let $p$ be an integer $\geq 2$. If $0 \leq a_n \leq p-1$, where $a_n$ is an integer then $\sum\limits_{n=0}^{\infty} \frac{a_n}{p^n}$ converges to some $x$ in $[0,1]$.
\end{lem}
\begin{proof}
    Since $0 \leq a_n \leq p-1$, we can replace all $a_n$ with $p-1$ and then we have$$ \sum\limits_{n=1}^{\infty} \frac{a_n}{p^n} \leq (p-1)\sum\limits_{n=1}^{\infty} \frac{1}{p^n} = 1$$
    Therefore, the sequence is bounded, and it is monotonic increasing this means it converges to some $x \leq 1$ and $x$ is already positive. Hence it converges tp some $x$ in $[0,1]$. 
\end{proof}
\begin{lem}
    Conversely, given any  $ 0 \leq x \leq 1, \exists~ a_n \in \mathbb{Z}$ and $0 \leq  a_n \leq p-1$ such that $x = \sum\limits_{n=0}^{\infty} \frac{a_n}{p^n}$.
\end{lem}
\begin{proof}
    Suppose we have $0 < x \leq 1$ and $a_1$ is the largest integer such that $ \frac{a_1}{p} < x \leq 1$. Since $x$ is bounded above by $1$, we have $ a_1 < p$ $\implies$ $a_1 \leq p-1$ since $a_1$ is an integer. Similarly, find $a_2$ such that $\frac{a_1}{p} + \frac{a_2}{p^2} < x$.This can be achieved by Archimedean property. Also, note that $a_2 \leq p-1$, since we have 
    \begin{align*}
        \frac{a_1}{p} + \frac{a_2}{p^2} &< x < 1\\
        \frac{a_1}{p} + \frac{a_2}{p^2} \leq \frac{p-1}{p} + \frac{a_2}{p^2} &< 1~~~~ (a_1 \leq p-1)\\
        1 - \frac{1}{p} + \frac{a_2}{p^2} &< 1\\
        \frac{a_2}{p^2} &< \frac{1}{p}\\
        a_2 &< p \\
        a_2 &\leq p-1
    \end{align*}
Inductively, we can define $a_n$ as the largest integer with $a_n \leq p-1$ such that $\sum\limits_{i=1}^{n} \frac{a_i}{p_i} < x$. Since $a_n < p$
TBD
\end{proof}
Suppose $\{a_n\}$ is the bounded sequence in $\mathbb{R}$, we define two sets: 
\begin{align*}
    s_n &:= inf\{a_n, a_{n+1}, \dots \}\\
    S_n &:= sup\{a_n, a_{n+1}, \dots \}
\end{align*}
Notice that $ \underset{k}{inf}(\{a_n\}) \leq s_n \leq S_n \leq \underset{k}{sup}(\{s_n\})$
\begin{defn}
    Limit superior and limit inferior: Let $a_n$ be the bounded sequence of real numbers then
    \begin{align*}
        \lim inf(a_n) &= \lim_{n\to \infty} s_n\\
        \lim sup(a_n) &= \lim_{n\to \infty} S_n
    \end{align*}
\end{defn}
Limit superior is the supremum of all subsequential limits of $\{a_n\}$. Similarly, limit inferior is the infimum of all subsequential limits of $\{a_n\}$.

Note that $s_n$ is the increasing sequence and $S_n$ is the decreasing sequence and they are bounded on both sides. Hence, we can also say that using monotone convergence theorem
\begin{align*}
    \lim inf(a_n) &= \lim_{n\to \infty} s_n = sup(s_n)\\
    \lim sup(a_n) &= \lim_{n\to \infty} S_n = inf(S_n)
\end{align*}
\begin{thm}
    A sequence $\{a_n\}$ is bounded above iff limsup $a_n < \infty$ (is finite).
\end{thm}
\begin{proof}
    TBD
\end{proof}
\begin{thm}
    A sequence $\{a_n\}$ is bounded below iff liminf $a_n < \infty$ (is finite).
\end{thm}
\begin{proof}
    TBD
\end{proof}
\begin{thm}
    Given any sequence $\{a_n\}$ there exists a subsequence $\{a_{n_k}\}$ such that $a_{n_k} \to limsup(a_n)$.
\end{thm}
\begin{proof}
    TBD
\end{proof}
\begin{thm}
    Given any sequence $\{a_n\}$ there exists a subsequence $\{a_{n_k}\}$ such that $a_{n_k} \to liminf(a_n)$.
\end{thm}
\begin{proof}
    TBD
\end{proof}
\hrule
\section*{Lecture 3}
\begin{defn}
    Finite Set: A set is finite if there exists a bijection between the set and the $\{1,2, \dots, n\}$ for some $n \in \mathbb{N}$.
\end{defn}
\begin{thm}
    $\mathbb{N}$ is an infinite set.
\end{thm}
\begin{proof}
    Negation of above definition would be a set is infinite if there does not exists a bijection between $\{1,2, \dots, n\}$ and $\mathbb{N}$. Suppose a function $f(\{1,2, \dots, n\}) \to \mathbb{N}$. \\
    We can have a natural number $f(1)+f(2)+\dots+f(n) > f(i)~~\forall i \in \mathbb{N}$ that does not have a preimage in $\{1,2, \dots, n\}$ hence the map is not bijective.
\end{proof}
\begin{thm}
    A set is infinite iff there exists a one-one map from $\mathbb{N}$ to the set.
\end{thm}
\begin{proof}
    $\implies$) For $1 \in \mathbb{N}$ there exist a image in $X$ say $f(1)$. Now, take $2 \in \mathbb{N}$ such that there exist a image $f(2) \in X\backslash \{1\}$. This means that $f(1) \neq f(2)$. Since for every $n \in \mathbb{N}$ we can have $f(n)$ in $X\backslash\{1,2, \dots, n-1\}$ as $X$ is also infinite. Hence, we have constructed a one-one map from $\mathbb{N} \to X$.\\
    ($\impliedby$ Since $\mathbb{N}$ is infinite and we have a one-one mapping from $\mathbb{N} \to X$ therefore for each $n \in \mathbb{N}$, we have only one $f(n) \in X$ and every $n \in \mathbb{N}$ has a image (it's a map). Hence, $X$ is infinite.
\end{proof}
\begin{defn}
    Equivalent or Equipotent set: Two sets are equivalent or equipotent if there exists bijection between $X$ and $Y$.
\end{defn}
For example: Finite sets are equivalent to $\{1,2, \dots, n\}$ for some fixed $n \in \mathbb{N}$.
\begin{defn}
    Countably infinite set: A infinite set $X$ is said to be countably infinite if there exists a bijection between $X$ and $\mathbb{N}$. 
\end{defn}
\begin{defn}
    Uncountably infinite set: A set is said to be uncountably infinite if it is not countably infinite set. 
\end{defn}
\textbf{Example:} 
\begin{enumerate}
    \item Countably infinite: bijective map between $\mathbb{Z} \to \mathbb{N}.$ The map will look line
    $$ 
    n\to \begin{cases}
            \frac{n}{2}& n\in even\\
            \frac{-(n+1)}{2}&~~ n \in odd    
         \end{cases}
    $$
    \item $\mathbb{N} \times \mathbb{N} \to \mathbb{N}$ is also equivalent to $\mathbb{N}$ i.e. countably infinite. Map will be $ (m \times n) \to 2^m(2n-1)$ $m$ and $n$ are unique and $m$ such that $2^m$ is the maximum multiple of $2$.
\end{enumerate}
\begin{thm}
    If $A$ is an infinite subset of $\mathbb{N}$ then there exists bijection between $A$ and $\mathbb{N}$.
\end{thm}
\begin{proof}
    TBD
\end{proof}
\begin{cor}
    Monotone Subsequence theorem: Any sequence $\{x_n\}$ of real numbers has a monotone subsequence
\end{cor}
\begin{proof}
    We define "peak" as any element $x_m$ is called a peak if $x_m\geq x_n$ for all $n > m$. There cases can be two possible cases
    \begin{enumerate}
        \item Infinite peaks: This means that there exists $m_i's$ say $\{m_1, m_2, \dots\}$ such that $x_{m_i} > x_n$ for all $n> m_i ,~\text{and for all}~i \in \mathbb{N}$. We can arrange $m_i's$ in increasing order $m_1 < m_2 < \dots$ and $x_{m_1} > x_{m_2} > x_{m_3} > \dots$ is a decreasing subsequence.
        \item Finite peaks: ($0$ or some $n \in \mathbb{N}$). Assume the peaks are $ \{x_{m_1}, x_{m_2} , \dots, x_{m_n}\}$, this means that there exists some $s_1 = m_n+1, s \in \mathbb{N}$ such that $x_{s_1}$ is not a peak. Therefore, there also exists some $s_2 > s_1$ such that $x_{s_1} < x_{s_2}$. Because of finite peaks $x_{s_2}$ is also not a peak, hence for some $s_3 > s_2$, we have $x_{s_2} < x_{s_3}$. Proceeding with induction, we have $x_{s_1} < x_{s_2} < x_{s_3} < \dots$ is a increasing sequence. 
    \end{enumerate}
\end{proof}
\begin{thm}
    $\mathbb{Q}$ is a countably infinite set.
\end{thm}
\begin{proof}
    TBD
\end{proof}
\begin{thm}
    Any interval in $\mathbb{R}$ is an uncountable set.
\end{thm}
\begin{proof}
    TBD
\end{proof}
\begin{defn}
    Suppose $X \neq \phi$. A partial order on $X$ is a relation $R$ on $X$ such that $R$ is
    \begin{enumerate}
        \item Reflexive.
        \item Anti-symmetric $\implies$ $aRb, bRa \implies a=b$
        \item Transitive.
    \end{enumerate} 
\end{defn}
\textbf{Examples:}
\begin{enumerate}
    \item $R = "\leq"$ is a partial order.
\end{enumerate}
$\mathcal{P(A)}$ is power set of $A$ and $X,Y \subset A$ then $X \leq Y$ iff $X \subseteq Y$.
\hrule
\section*{Lecture 4}
\begin{defn}
    Given $E \subset X$ where $X$ is partially order set, we say $E$ is totally ordered if any two elements of $E$ are comparable i.e. if $e_1,e_2 \in E$, then $e_1 \leq e_2$ or $e_2 \leq e_1$. Totally ordered $\equiv$ linearly order $\equiv$ chain.
\end{defn}
\begin{defn}
    Upper bound of $E$: An element is $x \in X$ is called upper bound of $E$ if for any $x' \in E$, we have $x' \leq x$. $x$ is called the maximal element if $x' \geq x \implies x' = x$.\\
    For maximal element, $x$ should be an upper bound for set $E$ and $x$ should belong to $E$.
\end{defn}
Let $X \neq \phi$. $\mathcal{F}$ is a collection of subsets of $X$ (element of $\mathcal{P(\text{X})}$). An element $F \in \mathcal{F}$ is a upper bound for a subfamily $\mathcal{F'}$ of $\mathcal{F}$ provided every member of $\mathcal{F'}$ is a subset of $F$.\\
$F$ will be the maximal element of $\mathcal{F}$ if it is not a proper subset (means not contained in) of any member in $\mathcal{F}$. 
\begin{lem}
    Zorn's lemma: Let $X$ be a partially order set. If every totally ordered subset of $X$ is bounded above then $X$ has a maximal element.
\end{lem}
\begin{defn}
    Cardinality of $X:$ Two sets $A$ and $B$ have same cardinality if there exists a bijection between them. Set $X$ has cardinal number $\alpha$ means that there exists a set $Y$ equivalent to $X$ with number of elements equal to $\alpha$.   
\end{defn}
If $\alpha$ and $\beta$ are cardinal numbers of set $X$ and set $Y$ such that $\alpha \leq \beta$ then there exists a one-one mapping from $X \to Y$.
\begin{thm}
    Cantor-Schroeder-Bernstein theorem: If there exists a one-one mapping from $X \to Y$ and $Y \to X$ then there exists a bijection between $X$ and $Y$.
\end{thm}
\subsection*{Limits of functions}
\begin{defn}
    Limit point: A point $a \in \mathbb{R}$ is called a limit point of a set $X \subseteq \mathbb{R}$ if for every neighbourhood $(a - \epsilon, a+ \epsilon), \epsilon>0$ there exists $x \in X$ such that $a \neq x$.  
\end{defn}
For a function $f$ defined on $X \subseteq \mathbb{R}$, $f$ converges to some $l$ means that for every $\epsilon > 0$ there exists a $\delta > 0$ such that
$$ |f(x) - l| < \epsilon ~\forall x \in X \ni |x - a| < \delta$$ 
Limit of functions as limit of sequences-
\begin{prop}
    Let $f:X\to \mathbb{R}$ and let $a$ be a limit of $X$. Then $lim_{x \to a} f(x) = l$ if and only if for every sequence $\{x_n\}_{n \geq 1}$ in $X$ that converges to $a$ and $x_n \neq a$ for all $n$, the sequence $\{f(x_n)\}_{n \geq 1}$ converges to $l$. 
\end{prop}
A function $f$ is continuous on $X$ if it is continuous in every point in $X$.
\begin{defn}
    Continuity of $f$: A function $f$ is said to be continuous at some point $x \in X$ if for every $\epsilon > 0$ there exists a $\delta > 0$ such that for all $y \in X$ and $|y - x| < \delta$, we have $|f(y) - f(x)| < \epsilon$.
\end{defn}
A function $f$ is continuous at a limit point $a \in X$ if and only if $f(a)$ is defined and $ lim_{x \to a} f(x) = f(a)$.
\begin{prop}
    Let $f$ be a real valued funtion defined on subset $X$ of $\mathbb{R}$ and $a \in X$ is the limit point of $X$. Then $f$ is continuous at $a$ if and only if for every sequence $\{x_n\}_{n \geq 1}$ that converges to $a$ and $x_n \neq a$ for every $n$, we have $lim f(x_n) = f(lim x_n) = f(a)$. Continuous function preserve convergence (maps convergent sequence into convergent sequences).  
\end{prop}
\begin{thm}
    Bolzano intermediate value theorem: Let $I$ be an interval and $f: I \to \mathbb{R}$, if $a,b \in I$ and $\alpha \in \mathbb{R}$ satisfies $f(a) < \alpha < f(b)$ or $f(a) > \alpha > f(b)$ then there exists a point $c \in I$ between $a$ and $b$ such that $f(c) = \alpha$.
\end{thm}
\subsection*{Sequences of functions}
Let $X \subseteq \mathbb{R}$. If for every $n = 1,2, \dots$, we assigned a real valued function $f_n$ defined on $X$ then $\{f_n\}_{n \geq 1}$ is called sequence of  functions.\\
\textbf{Point-wise convergence}\\
A sequence is called point-wise convergent if for each $x \in X \subseteq \mathbb{R}$, the sequence $f_1(x), f_2(x), \dots$ of real numbers is convergent. \\
\textbf{Point-wise limit} \\
A function defined on $X$ is called a point-wise limit if for every $\epsilon > 0$ there exist $n_0 \in \mathbb{Z}$ depending on $x$ and $\epsilon$ such that for all $n \geq n_0$, we have $|f(x) - f_n(x)| < \epsilon$.

A series $\sum\limits_{n = 1}^{\infty} x_n$ of real numbers converges to $x \in \mathbb{R}$ if the sequence of partial sums $\{s_n\}_{n\geq 1}$ converges to $x$ where $s_n = \sum\limits_{k = 1}^n x_k$ ($n_{th}$ partial sum).\\
The limit $x = lims_n = lim_{n \to \infty}\sum\limits_{k = 1}^n x_k$ is called the sum of the series. In the case of functions, if for every $x \in X$, $\sum\limits_{n = 1}^{\infty} f(x_n)$ converges and if we define $f(x) = \sum\limits_{n = 1}^{\infty} f(x_n)$, the $f(x)$ is called the sum of the series $\sum\limits_{n = 1}^{\infty} f_n$.\\
\textbf{Uniform convergence}\\
A sequence of functions $\{f_n\}_{n \geq 1}$ defined on set $X \subseteq \mathbb{R}$ is said to be uniformly convergent on $X$ to $f$ if for every $\epsilon > 0$ there exists a $n_0 \in \mathbb{Z}$ (depending on $\epsilon$ only) such that for all $n \geq n_0$, we have $|f(x) - f_n(x)| < \epsilon$ for all $x \in X$.\\
Thus, every uniformly convergent sequence is pointwise convergent but converse is not true always.\\
\textbf{Uniform Convergence for series of functions}\\
The series of functions $\sum\limits_{n=1}^{\infty}$ converges uniformly on $X$ if the sequence $\{s_n\}_{n \geq 1}$ partial sum of functions, where $s_n(x) = \sum\limits_{k=1}^n f_k(x), x \in X$ converges uniformly on $X$.
\textbf{Cauchy Criterion of Uniform convergence}\\
The sequence of functions $\{f_n\}_{n\geq 1}$ defined on $X \in \mathbb{R}$ converges uniformly on $X$ if and only if given $\epsilon > 0$ there exists $n_0$ such that for all $x \in X$, and all $m \geq n_0,n\geq n_0$ we have $|f_m(x) -f_n(x)| < \epsilon$.
\begin{prop}
    Suppose $\{f_n\}_{n\geq 1}$ is sequence of continuous functions defined on $X$ and it converges uniformly to $f$ then $f$ is continous.
\end{prop}
\subsection*{Basic Topology in $\mathbb{R}$}
\begin{defn}
    Open sets in $\mathbb{R}$: A subset $G$ of $\mathbb{R}$ is said to be open if for every $x \in G$, there is a neighbourhood $(x -\epsilon, x+\epsilon), \epsilon > 0$ that is contained in $G$.
\end{defn}
\begin{defn}
    Open cover: A open cover of $X$ is a collection of $C = \{G_{\alpha} \mid \alpha \in I\}$ of open sets in $\mathbb{R}$ whose union contains the set $X$, $$ X \subseteq \bigcup\limits_{\alpha} G_{\alpha} $$ where $I$ is some indexing set.
\end{defn}
\begin{defn}
    Sub-cover: If $C'$ is a subcollection of $C$ such that the union of sets in $C'$ also contains the set $X$ then $C'$ is called subcover from $C$ of $X$. If the number of sets in $C'$ is finite then we call it a finite subcover. 
\end{defn}
\begin{defn}
    Compact set: A subset $X$ of $\mathbb{R}$ is said to be compact if every open cover of $X$ has a finite sub-cover. 
\end{defn}
\begin{prop}
    (Heine-Borel Theorem) Let $X$ be a set of real numbers. Then the following statements are equivalent:
    \begin{enumerate}
        \item $X$ is closed and bounded.
        \item $X$ is compact (every open cover has a finite subcover).
        \item Every infinite subset of $X$ has a limit point in $X$.
    \end{enumerate}
\end{prop}
\begin{proof}
    $1\implies 2$) Suppose $X = [a,b]$ a infinite interval which is closed and bounded. We define a set $S$
    $$ S = \{x \in [a,b] \mid \exists n \in I \ni [a,x] \subseteq \bigcup\limits_{i = 1}^{i = n} O_i\}$$

    $S$ is non-empty as we can always choose $n$ such that $S$ contains atleast $a$. Also, $b$ is the upper bound of the set $S$ and then it should possess the least upper bound property. Let $\lambda = sup(S)$.

    Basically, $S$ is a collection of all $x$ such that the interval $[a, x]$ has a finite subcover. Hence, we need to show that $\lambda$ is largest possible element of $S$ (i.e. $\lambda \in S$) for which $[a, \lambda]$ has a finite subcover and $\lambda = b$.

    Since, $\lambda = sup(S)$ and $\lambda < b$ hence, $\lambda \in X$. This means that there exists atleast one open set $O_{\beta}$ that contains $\lambda$ for  some $\beta \in I$. Hence, for some $\epsilon > 0$, the neighbourhood $(\lambda - \epsilon, \lambda + \epsilon) \in O_{\beta}$. Since, $\lambda - \epsilon$ is not the supremum of set $S$. Therefore, there exists a element $x \in S$ such that $\lambda > x > \lambda - \epsilon$ $\implies$ $x \in O_{\beta}$. Also, $x \in S$ hence, by definition of $S$ the finite subcovers of $[a, x]$ are $\{ O_1, O_2, \dots, O_{n}\}$ for some fixed $n \in I$. Hence, adding $O_{\beta}$ to this set $\{ O_{beta}, O_1, O_2, \dots, O_{n}\}$ is also finite and it is subcover of set $[a, \lambda]$ $\implies$ $\lambda \in S$.

    For proving $\lambda = b$, we have $\lambda \leq b$. Suppose for contradiction that $\lambda < b$ $\implies$ $b - \lambda > 0$. Define $y := \lambda + \frac{1}{2}\min(\epsilon, b-\lambda)$. This implies that belongs $y$ to neighbourhood of $\lambda$ $\implies$ $ y \in O_{\beta}$ and $y \in \bigcup\{ O_{\beta}, O_1, O_2, \dots, O_{n}\}$ $\implies y \leq \lambda$. But by definition $y > \lambda$. Hence, contradiction. Therefore, $y = b$ is the only possibility. This means that $X:= [a, b]$ has finite subcover hence $X$ is compact.
\end{proof}
\begin{proof}
    $2 \implies 3$) Let a infinite subset $G$ of $X$ where $X$ is compact set. Suppose for contradiction $G$ does not have any limit point in $X$ TBD.
\end{proof}
\begin{prop}
    Let $f$ be real-valued continuous function defined on closed and bounded interval $I = [a, b]$. Then $f$ is bounded on $I$ and it assumes its maximum and minimum values in the interval $I$ i.e. there are points $x_1, x_2 \in I$ such that $f(X_1) \leq f(x)  \leq f(x_2)$ for all $x \in X$.
\end{prop}

\paragraph*{Uniformly Continuous function}
Let $f$ be real-valued continuous function defined on $X$. Then $f$ is said to be uniformly continuous if given $\epsilon > 0$, there is a $\delta > 0$ such that for all $x,y \in X$ with $|x - y| < \delta$, we have $|f(x) - f(y)| < \epsilon$.
\begin{prop}
    If a real-valued function $f$ is continous on a closed and bounded interval $I$ then $f$ is uniformly continuous on $I$.
\end{prop}
\subsection*{Sequences in $\mathbb{R}$}
\begin{enumerate}
    \item A convergent sequence in $\mathbb{Q}$ is a Cauchy sequence in $\mathbb{Q}$.
    \item A Cauchy sequence in $\mathbb{Q}$ is bounded; in particular, every convergent sequence in $\mathbb{Q}$ is bounded.
    \item Limit of a sequence say $\{x_n\}_{n \geq 1}$ is unique.
\end{enumerate}
\textbf{Let $F_Q$ denote the set of all cauchy sequence in $\mathbb{Q}$.}
\begin{defn}
    A sequence $\{x_n\}_{n \geq 1}$ in $F_Q$ is said to be equivalent to a sequence $\{y_n\}_{n \geq 1}$ in $F_Q$ if and only if $\lim_{n \to \infty}|x_n - y_n| = 0$. Notation of equivalence is $\{x_n\}_{n \geq 1} \sim \{y_n\}_{n \geq 1}$.
\end{defn}
\begin{prop}
    If $\{x_n\}_{n\geq 1} \in F_Q$ then $\lim_{n \to \infty} x_n = x$ if and only if $\{x_n\} \sim \{x\}$, where $\{x\}$ denotes the constant sequence with each term is equal to $x$. 
\end{prop}
\begin{proof}
    Follows from definition of equivalence relation.\\
    Since, $\lim_{n \to \infty} |x_n - x| = 0$ (from definition) $\implies$ $\lim_{n \to \infty} x_n =  \lim_{n \to \infty} x = x$. Converse, is similar. 
\end{proof}
\begin{prop}
    If $\{x_n\}$ and $\{y_n\}$ are in $F_Q$ then so the sequencs $\{x_n + y_n\}$ and $\{x_ny_n\}$.
\end{prop}
\begin{proof}
    Apply cauchy criterion for each sequence and choose $\epsilon$ to be $\frac{\epsilon}{2}$. Then try to find $\epsilon$ bound for sequences $\{x_n + y_n\}$ and $\{x_ny_n\}$.
\end{proof}
\begin{prop}
    If $\{x_n\}$, $\{y_n\}$, $\{x'_n\}$ and $\{y'_n\}$ are in $F_Q$ and $\{x_n\}\sim \{x'_n\}$, $\{y_n\}\sim \{y'_n\}$ then $\{x_n + x'_n\}\sim \{y_n + y'_n\}$ and $\{x_ny_n\} \sim \{x'_ny'_n\}$.
\end{prop}
\begin{proof}
    For $\{x_n\}\sim \{x'_n\}$, $\{y_n\}\sim \{y'_n\}$ follows from writing modulus inequality $|a| - |b| \leq |a-b| \leq |a+b| \leq |a| + |b|$ and then applying sandwich theorem.

    For $\{x_ny_n\} \sim \{x'_ny'_n\}$, we know that cauchy sequence in $\mathbb{Q}$ are bounded. Hence, there exists a rational $K_1,K_2$ such that $|x_n| \leq K_1$ and $|y'_n| \leq K_2$ for all $n$. We can write 
    $$ |x_n - x'_n| < \frac{\epsilon}{2K_1}~\text{and}~|y_n - y'_n| < \frac{\epsilon}{2K_2}$$
    \begin{align*}
        |x_ny_n - x'_ny'_n| &= |x_ny_n - x_ny'_n + x_ny'_n - x'_ny'_n|\\
        &= |x_n(y_n - y'_n) + y'_n(x_n - x'_n)|\\
        &\leq |x_n||(y_n - y'_n) | + |y'_n||(x_n - x'_n)|\\
        &< K_1\left(\frac{\epsilon}{2K_1}\right) + K_2\left(\frac{\epsilon}{2K_2}\right)\\
        &< \frac{\epsilon}{2} + \frac{\epsilon}{2} < \epsilon
    \end{align*}
\end{proof}
\subsection*{Inequalities}
\begin{prop}
    The function $f(x) = \frac{x}{1+x}, x \geq 0,$ is monotonically increasing. 
\end{prop}
\begin{proof}
    For some $x, y \geq 0$ and $x > y$, we have $\frac{1}{1+x} < \frac{1}{1+y}$ and $1-\frac{1}{1+x} > 1-\frac{1}{1+y}$ $\implies$ $ \frac{x}{1+x} > \frac{y}{1+y}$. 
\end{proof}
\begin{thm}
    For any two real numbers $x$ and $y$, this inequality holds
    $$ \frac{|x+y|}{1+|x+y|} \leq \frac{|x|}{1+|x|} + \frac{|y|}{1+|y|}$$
\end{thm}
\begin{proof}
    Using previous proposition, we can say if $|x+y| \leq |x| + |y|$ and the sequence $\frac{x}{1+x}, x \geq 0$ is monotonically increasing then
    $$ \frac{|x+y|}{1+|x+y|} \leq \frac{|x| + |y|}{1+ |x| + |y|} = \frac{|x|}{1+|x| + |y|} + \frac{|y|}{1+|x| + |y|} \leq \frac{|x|}{1+|x|} + \frac{|y|}{1+|y|}$$ 
\end{proof}
\begin{prop}
    Generalised AM-GM inequality: If $a > 0$ and $b > 0$ and if $0 \leq \lambda \leq 1$ is fixed, then
    $$ a^{\lambda}b^{1-\lambda} \leq \lambda a + (1-\lambda)b$$ 
\end{prop}
\begin{proof}
    Since, $ y = ln(x)$ is concave, then
    \begin{align*}
        ln(\lambda a + (1-\lambda)b) &\geq \lambda ln(a) + (1-\lambda)ln(b)\\
        ln(\lambda a + (1-\lambda)b) &\geq ln(a^{\lambda} b^{1-\lambda})
    \end{align*}
Since $e^x$ is increasing function, we have
$$ a^{\lambda} b^{1-\lambda} \leq \lambda a + (1-\lambda)b$$
\end{proof}
\begin{rem}
    When $x \geq 0, y \geq 0$ and $p > 1$ and $\frac{1}{p} + \frac{1}{q} = 1$, we have 
    $$ xy \leq \frac{1}{p}x^p + \frac{1}{q}y^q$$
\end{rem}
\begin{proof}
    Replace $ a = x^p$, $b = y^q$ and take $\lambda = \frac{1}{p} \implies 1-\lambda = 1- \frac{1}{p} = \frac{1}{q}$
\end{proof}
\begin{thm}
    (Holder's inequality) Let $x_i \geq 0$ and $y_i \geq 0$ for $i = 1,2, \dots, n,$ and suppose that $p > 1$ and $q > 1$ are such that $\frac{1}{p} + \frac{1}{q} = 1$. Then
    $$ \sum\limits_{i=1}^n x_iy_i \leq \left(\sum\limits_{i=1}^nx_i^p\right)^{1/p} \left(\sum\limits_{i=1}^ny_i^q\right)^{1/q}$$
    In the special case when $p = q = 2$, the above inequality reduces to 
    $$ \sum\limits_{i=1}^n x_iy_i \leq \left(\sum\limits_{i=1}^nx_i^2\right)^{1/2} \left(\sum\limits_{i=1}^ny_i^2\right)^{1/2}$$
    This is called \textbf{Cauchy-Schwarz inequality.}
\end{thm}
\begin{proof}
    For the case of $x_i = 0$ and $y_i = 0$, it is trivially true.\\
    For the case of $x_i > 0$ and $y_i > 0$, we can write the given inequality as 
    $$ \sum\limits_{i = 1}^{n} \left(\frac{x_i}{\left(\sum\limits_{i = 1}^{n} x_i^p\right)^{1/p}} \frac{y_i}{\left(\sum\limits_{i = 1}^{n} y_i^q\right)^{1/q}} \right)\leq 1 $$
    Replace $ x_i' = \frac{x_i}{\left(\sum\limits_{i = 1}^{n} x_i^p\right)^2}$ and $y_i' = \frac{y_i}{\left(\sum\limits_{i = 1}^{n} y_i^q\right)^2} $, also $x_i' \geq 0$ and $y_i' \geq 0$, we get $ \sum\limits_{n = 1}^{n} x_i'y_i' \leq 1$. \\
    Now,  apply the Youngs inequality for $i = 1, 2, \dots, n$ and sum them up to get
    $$ \sum\limits_{i = 1}^{n} x_i'y_i' \leq \frac{\sum x^p}{p} + \frac{\sum y^q}{q}$$
    Since, $x_i > 0$ and $y_i > 0$ hence $\sum\limits_{n=1}^{n}x'^p \neq 0 \neq \sum\limits_{n=1}^{n}y'^q $ and it is equivalent to prove it for (some constant)
    $$ \sum\limits_{n=1}^{n}x'^p = 1 = \sum\limits_{n=1}^{n}y'^q $$
    Hence, we have
    $$ \sum\limits_{i = 1}^{n} x_i'y_i' \leq \frac{1}{p} + \frac{1}{q} = 1$$
    This proves the required inequality.
\end{proof}
\begin{thm}
    (Minkowski's inequality) Let $x_i \geq 0$ and $y_i \geq 0$ for $ i = 1,2,\dots, n$ and suppose that $p \geq 1$. Then
    $$ \left(\sum\limits_{i =1}^n(x_i + y_i)^p\right)^{1/p} \leq \left(\sum\limits_{i =1}^nx_i^p\right)^{1/p} + \left(\sum\limits_{i =1}^ny_i^q\right)^{1/q}$$  
\end{thm}
\begin{proof}
    If $p = 1$, it is trivially true. So, assume $p > 1$, we have
    $$ \sum\limits_{i=1}^n(x_i + y_i)^p = \sum\limits_{i=1}^{n} x_i(x_i + y_i)^{p-1} + \sum\limits_{i = 1}^n y_i(x_i + y_i)^{p-1}$$
    Apply Holder's inequality for both terms on RHS
    \begin{align*}
        \sum\limits_{i=1}^n(x_i + y_i)^p &\leq \left(\sum\limits_{i=1}^n x_i^p \right)^{1/p} \left(\sum\limits_{i=1}^n (x_i + y_i)^{(p-1)q}\right)^{1/q} \\
        &+ \left(\sum\limits_{i=1}^n y_i^p \right)^{1/p} \left(\sum\limits_{i=1}^n (x_i + y_i)^{(p-1)q}\right)^{1/q}\\
        &\leq \left[\left(\sum\limits_{i=1}^n x_i^p\right)^{1/p} + \left(\sum\limits_{i=1}^n y_i^p\right)^{1/p}\right] \left(\sum\limits_{i=1}^n(x_i + y_i)^p\right)^{1/q}
    \end{align*}
    Divide both sides by $\left(\sum\limits_{i=1}^n(x_i + y_i)^p\right)^{1/q}$ as it is $\neq 0$ and we will get the required inequality. For the case of $\left(\sum\limits_{i=1}^n(x_i + y_i)^p\right)^{1/q} = 0$, proof is self-evident. 
\end{proof}
\begin{thm}
    (Minkowski Inequality for infinite sums) Suppose that $p \geq 1$ and let $\{x_n\}_{n \geq 1}$ and $\{y_n\}_{n \geq 1}$ be sequences of non-negative terms such that $\sum\limits_{n=1}^{\infty} x_n^p$ and $\sum\limits_{n=1}^{\infty} y_n^p$ are convergent. Then $\sum\limits_{n=1}^{\infty} (x_n + y_n)^p$ is convergent. Moreover,
    $$ \left(\sum\limits_{n =1}^{\infty}(x_i + y_i)^p\right)^{1/p} \leq \left(\sum\limits_{n =1}^{\infty}x_i^p\right)^{1/p} + \left(\sum\limits_{n =1}^{\infty}y_i^q\right)^{1/q} $$
\end{thm}
\begin{proof}
    For any positive integer $m$, we can use Minkowski inequality for finite sums 
    \begin{align*}
        \left(\sum\limits_{n =1}^m(x_i + y_i)^p\right)^{1/p} &\leq \left(\sum\limits_{n =1}^mx_i^p\right)^{1/p} + \left(\sum\limits_{n =1}^my_i^q\right)^{1/q}\\
        &\leq \left(\sum\limits_{n =1}^{\infty}x_i^p\right)^{1/p} + \left(\sum\limits_{n =1}^{\infty}y_i^q\right)^{1/q}
    \end{align*}
    Since, the sequence $\{(\sum\limits_{n =1}^m(x_i + y_i)^p)^{1/p}\}$ is increasing sequence(w.r.t m) of nonnegative real numbers and is bounded above by the sum in above equation. Hence, it is convergent and as $m \to \infty$ the limit is also bounded above by the sum. This proves the inequality.
\end{proof}
\begin{thm}
    Let $p > 1$. For $ a \geq 0$ and $ b \geq 0$, we have
    $$ (a + b)^p \leq 2^{p-1}(a^p + b^p)$$
\end{thm}
\begin{proof}
    TBD
\end{proof}
\subsection*{Metric Spaces}
\begin{defn}
    Let $X \neq \phi$. A metric on $X$ is a function $d:X \times X \to [0, \infty)$ such that 
    \begin{enumerate}
        \item d(x,y) = 0 $\implies x = y$
        \item d(x,y) = d(y,x)
        \item d(x,y) $\leq$ d(x,z) + d(z,y) for any $x,y,z \in X$.
    \end{enumerate}
\end{defn}
The metric space in which every cauchy sequence converges is called "complete" metric space.
\subsection*{Examples of metric spaces}
\begin{enumerate}
    \item \textbf{The space of all bounded sequences.} Let $X$ be the set of all infinite sequences of numbers that are bounded $\{x_i\}_{i \geq 1}$ such that $sup|\{x_i\}| < \infty$. The metric on $X$ is defined as $d(x,y)= sup_i|x_i - y_i| ~(\leq sup_i|x_i - z_i| + sup_i|y_i - z_i|).$
    \item \textbf{The space $l_p$.} Let $X$ be the set of all sequences $x = \{x\}_{i \geq 1}$ such that 
    $$ \left(\sum\limits_{i = 1}^{\infty} |x_p|^p\right)^{1/p} < \infty,~~~~p \geq 1$$
    If $\{x\}_{n \geq 1}$ and $\{y\}_{n \geq 1}$ are two sequences belong to $X$ then we define the metric as
    $$ d(x,y) = \left(\sum\limits_{i = 1}^{\infty} |x_i - y_i|^p\right)^{1/p}$$
    Minkowski inequality for infinite sequences makes it a suitable metric.
    \item \textbf{The space of bounded functions.} Let $S$ be any nonempty set and $\mathcal{B(S)}$ denote the set of all real or complex-valued functions on $S$, each of which is bounded. i.e. 
    $$ \sup\limits_{x \in S} |f(x)| < \infty$$
    It can also be shown that $\sup\limits_{x \in S} |f(x) - g(x)| < \infty$. We defined metric as $d(f,g) = \sup\limits_{x \in S}|f(x) - g(x)|,~~~~ f,g \in \mathcal{B(S)}$. The metric $d$ is called \textbf{uniform metric} or \textbf{supremum metric}.
    \item \textbf{The space of continuous functions.} Let $X$ be the set of all continuous functions defined on $[a,b]$, an interval in $\mathbb{R}$. For $f,g \in X$, define
    $$ d(f,g) = \sup\limits_{x\in [a,b]} |f(x) - g(x)|$$
    We can also defined another metric on the set $X$ as
    $$ d(f,g) = \int_a^b |f(x) - g(x)| dx$$
    Assume $h(x) \in X = C[a,b]$, we have
    \begin{align*}
        d(f,g) &= \int_a^b |f(x) - h(x) + h(x) - g(x)| dx\\
        &\leq \int_a^b \left(|f(x) - h(x)| + |h(x) - g(x)| \right)dx\\
        &\leq d(f,h) + d(h,g)
    \end{align*}
    \item \textbf{Metric on extended real line.} Let $X = \mathbb{R} \cup \{\infty\} \cup \{-\infty\}$. Define $f: X \to \mathbb{R}$ by the rule
    $$ f(x) = \begin{cases}
        \frac{x}{1+|x|} & \text{if } -\infty < x < \infty \\
        1 & \text{if } x = \infty\\
        -1 & \text{if } x = -\infty
    \end{cases}
    $$
    We define metric using $f$ as 
    $$ d(x,y) = |f(x) - f(y)|, ~~ x,y \in X$$
    \begin{proof}
        TBD
    \end{proof}
    
\end{enumerate}
\end{document}

